network_analysis

### 2020-07-13 analyses/beta_diversity.Rmd
update

### Load network and survey data and generate site.year and site.month.year webs
```{r}
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  group_by(site, date) %>%
  #mutate(network.size = n()) %>%
  arrange(site, date)

@@

web_site <- net %>%
  group_by(site, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% 
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web
```

### Tabulate webs
S	= beta_S, the dissimilarity in species composition
OS = beta_OS, the dissimilarity (component) explained by "rewiring" among shared species
WN = beta_WN, the dissimilarity between the two networks
ST = beta_ST, the dissimilarity (component) explained by difference in species community composition 
```{r}
array_site <- webs2array(web_site)
array_site.year <- webs2array(web_site.year)

beta_site <- betalinkr_multi(array_site, partition.st = TRUE, partition.rr = TRUE) %>%
  as_tibble() %>%
  rename(site1 = i) %>%
  rename(site2 = j)%>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  left_join(site_transects, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
  left_join(site_transects, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         transects.diff = abs(transects1 - transects2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, site2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, 
                           lat1, lon2, lat2, transects1, transects2, transects.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

@@

beta_site.year <- betalinkr_multi(array_site.year, partition.st = TRUE, partition.rr = TRUE, binary = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "year1"), sep = "_") %>%
  separate(j, into = c("site2", "year2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  left_join(site.year_transects, by = c("site1" = "site", "year1" = "year")) %>%
  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
  left_join(site.year_transects, by = c("site2" = "site", "year2" = "year")) %>%
  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         transects.diff = abs(transects1 - transects2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, site2, year2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2,
                           transects1, transects2, transects.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))
```

>>>>

### Load network and survey data and generate site.year and site.month.year webs
```{r}
net <- read_csv("../processed_data/network.csv")%>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  group_by(site, date) %>%
  arrange(site, date)

@@

web_site <- net %>%
  group_by(site, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% 
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site.date.2010 <- net %>%
  filter(year == 2010) %>%
  group_by(site.year.day, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site.date.2011 <- net %>%
  filter(year == 2011) %>%
  group_by(site.year.day, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site.date.2012 <- net %>%
  filter(year == 2012) %>%
  group_by(site.year.day, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

array_site <- webs2array(web_site)
array_site.year <- webs2array(web_site.year)
array_site.date.2010 <- webs2array(web_site.date.2010)
array_site.date.2011 <- webs2array(web_site.date.2011)
array_site.date.2012 <- webs2array(web_site.date.2012)
```

### Tabulate webs
S	= beta_S, the dissimilarity in species composition
OS = beta_OS, the dissimilarity (component) explained by "rewiring" among shared species
WN = beta_WN, the dissimilarity between the two networks
ST = beta_ST, the dissimilarity (component) explained by difference in species community composition 
```{r}
beta_site <- betalinkr_multi(array_site, partition.st = TRUE, partition.rr = TRUE) %>%
  as_tibble() %>%
  rename(site1 = i) %>%
  rename(site2 = j)%>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  left_join(site_transects, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
  left_join(site_transects, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         transects.diff = abs(transects1 - transects2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, site2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, 
                           lat1, lon2, lat2, transects1, transects2, transects.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

@@

beta_site.year <- betalinkr_multi(array_site.year, partition.st = TRUE, partition.rr = TRUE, binary = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "year1"), sep = "_") %>%
  separate(j, into = c("site2", "year2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  left_join(site.year_transects, by = c("site1" = "site", "year1" = "year")) %>%
  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
  left_join(site.year_transects, by = c("site2" = "site", "year2" = "year")) %>%
  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         transects.diff = abs(transects1 - transects2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, site2, year2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2,
                           transects1, transects2, transects.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

### By individual transect

beta_site.date.2010 <- betalinkr_multi(array_site.date.2010, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "date1"), sep = "_") %>%
  separate(j, into = c("site2", "date2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site"))%>%
  dplyr::select(site1, date1, site2, date2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, date1, site2, date2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, date1, site2, date2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

beta_site.date.2011 <- betalinkr_multi(array_site.date.2011, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE)

beta_site.date.2012 <- betalinkr_multi(array_site.date.2012, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE)
```

### 2020-07-13 analyses/network_analysis.Rmd
update

    out <- bind_rows(out, temp)
  } 
  out <- out %>%
    gather(metric, value, -c(sp, site_year, iteration)) %>% # get mean value across iterations for each metric
    group_by(sp, site_year, metric) %>%
    summarize(value = mean(value)) %>%
    separate(site_year, c("site", "year"), sep = "_") %>%
    left_join(site_data) %>%
    left_join(bb_traits, by = c("sp" = "bb.sp"))
}

@@

# Yearly subsets
net_2012 <- filter(net, year == "2012")
net_2011 <- filter(net, year == "2011")
net_2010 <- filter(net, year == "2010")
```

@@

#### Species-level metric tables
```{r, message=FALSE}
### Species level
#### BBs
spmet_2012_BB <- boot_net_sp_BB(net_2012, 9, 40) %>%
  mutate(year = factor(rep("2012", n())))
spmet_2011_BB <- boot_net_sp_BB(net_2011, 4, 40) %>%
  mutate(year = factor(rep("2011", n())))  
spmet_2010_BB <- boot_net_sp_BB(net_2010, 3, 40) %>%
  mutate(year = factor(rep("2010", n())))

spmet_BB <- boot_net_sp_BB(net, 5, 50) %>%
  ungroup() %>%
  mutate(year = factor(year))

spmet_mean_BB <- spmet_BB %>%
  group_by(site, year, elev.mean, metric) %>%
  summarize(value = mean(value, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(year = factor(year))

spmet_mean_BB_2012 <- spmet_2012_BB %>%
  group_by(site, elev.mean, metric) %>%
  summarize(value = mean(value, na.rm = TRUE))

ggplot(spmet_mean_BB, aes(elev.mean, value)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(spmet_2012_BB, aes(elev.mean, value)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(spmet_mean_BB_2012, aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

spmet_BB_gam <- bam(log(value + 0.01) ~ year + 
                      s(elev.mean, bs = "tp", k = 10, m = 2) +
                      s(elev.mean, by = factor(sp), k = 20, m = 1),
                    data = spmet_BB,
                    method = "fREML",
                    discrete = TRUE) %>% getViz()

check.gamViz(spmet_BB_gam)
summary(spmet_BB_gam)
print(plot(spmet_BB_gam, allTerms = TRUE), pages = 1)

spmet_BB <- bind_rows(spmet_2012_BB, spmet_2011_BB, spmet_2010_BB)
```

>>>>

    out <- bind_rows(out, temp)
  } 
  out <- out %>%
    gather(metric, value, -c(sp, site_year, iteration)) %>% # get mean value across iterations for each metric
    group_by(sp, site_year, metric) %>%
    summarize(value = mean(value)) %>%
    separate(site_year, c("site", "year"), sep = "_") %>%
    left_join(site_data) 
  # %>%
  #   left_join(bb_traits, by = c("sp" = "bb.sp"))
}

@@

# Yearly subsets
net_2012 <- filter(net, year == "2012")
net_2011 <- filter(net, year == "2011")
net_2010 <- filter(net, year == "2010")

### Per-transect BB abundance: species-level
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # group by bb.sp*transect (and other variables we want to retain)
  summarize(bb.abund = n()) %>% # get bb.sp*transect abundance
  ungroup %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>%
  mutate(year = factor(year(date))) %>%
  mutate(yday = yday(date),
         sp = factor(bb.sp),
         site = factor(site)) %>%
  group_by(site, year, sp) %>%
  summarize(bb.total.abund = sum(bb.abund))
```

@@

#### Species-level metric tables
```{r, message=FALSE}
### Species level
#### BBs
spmet_2012 <- boot_net_sp(net_2012, 9, 40) %>%
  mutate(year = factor(rep("2012", n())))
spmet_2011 <- boot_net_sp(net_2011, 4, 40) %>%
  mutate(year = factor(rep("2011", n())))  
spmet_2010 <- boot_net_sp(net_2010, 3, 40) %>%
  mutate(year = factor(rep("2010", n())))

spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010) %>%
  left_join(bb_abund_sp) %>%
  mutate(sp = factor(sp),
         management = factor(management),
         site = factor(site))

ggplot(filter(spmet, bb.total.abund >= 20 & sp %in% c("bss", "pasc", "prat", "wurf", "soro")), aes(elev.mean, value, color = year)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~sp, scales = "free")

ggplot(filter(spmet, bb.total.abund >= 20), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~year)

specialization_GAM.2010 <- bam(value ~ 
                            s(elev.mean, by = sp) +
                            #s(bb.total.abund, by = sp) +
                            s(site, bs = "re") +
                            s(sp, bs = "re"),
                   data = filter(spmet, year == 2010),
                   discrete = TRUE,
                   select = TRUE,
                   method = "fREML") %>% getViz()

check.gamViz(specialization_GAM.2010)
specialization_GAM.2010.sum <- summary(specialization_GAM.2010)
print(plot(specialization_GAM.2010, allTerms = TRUE), pages = 1)

spmet_mean <- spmet %>%
  group_by(site, year, elev.mean, metric) %>%
  summarize(value = weighted.mean(value, bb.total.abund)) %>%
  ungroup()

ggplot(spmet_mean, aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~year)

```
